---
title: "Leinenger_2018_JEP:LMC"
author: "Mallorie Leinenger"
date: "June 28, 2018"
output: 
  github_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/TimeCoursePhonCoding/Revision2_Analyses")
```

# Leinenger M. (in press) Survival analyses reveal how early phonological processing affects eye movements during reading. *Journal of Experimental Psychology: Learning, Memory, and Cognition* 

## Overview
These data come from four experiments investigating the time course of phonological coding during reading. Participants read single-sentence stimuli (see example stimuli below) while their eye movement behavior was recorded. Sentences contained either a correct target word that was moderately predictable in context (e.g., *beach*, *sleep*), a contextually inappropriate (non-)word that shared phonology with the correct target word (e.g., *beech*, *sleap*), or a contextually inappropriate (non-)word that was match on orthographic overlap, but did not share phonology (e.g., *bench*, *slerp*). 

- The surfers traveled to the world-famous *beach/beech/bench* where the waves were very large. (Exps. 1 & 3)
- Because of his insomnia, Caleb couldn’t *sleep/sleap/slerp* even though he was tired. (Exps. 2 & 4)

In Experiments 1 & 2 participants directly fixated each type of target word, and in Experiments 3 & 4 the phonologically related and orthographic control (non-)words were presented as parafoveal previews using the invisible boundary display-change paradigm (Rayner, 1975). In addition to means analyses, survival analyses were conducted (following the method outlined in Reingold & Sheridan, 2014) to determine the earliest observable influence of phonological coding on the eye movement record.

## Abstract
Numerous studies have provided evidence that readers generate phonological codes while reading. However, a central question in much of this research has been how early these codes are generated. Answering this question has implications for the roles that phonological coding might play for skilled readers, especially whether phonological codes affect the identification of most words, which can only be the case if these codes are generated rapidly. To investigate the time course of phonological coding during silent reading, the present series of experiments examined survival analyses of first-fixation durations on phonologically related (homophones, pseudohomophones) and orthographic control (orthographically matched words and non-words) stimuli that were either embedded in sentences in place of correct targets (Experiments 1 and 2) or presented as parafoveal previews for correct targets using the boundary paradigm (Experiments 3 and 4). Survival analyses revealed a discernible difference between processing the phonologically related versus the orthographic control items by as early as 160 ms from the start of fixation on average (160–173 ms across experiments). Because only approximately 18% of first fixation durations were shorter than these mean estimates and follow-up tests revealed that earlier divergence point estimates were associated with shorter gaze durations (e.g., more rapid word identification), results suggest that skilled readers rapidly generate phonological codes during normal, silent reading and that these codes may affect the identification of most words.

```{r load packages, echo = FALSE, message=FALSE, warning=FALSE}
## load required packages
library(tidyverse)
library(lme4)
library(psych)
library(plotrix)
library(reshape2)
library(RTsurvival)
library(survival)
```

## Experiment 1
Data for 48 participants (**subject**) reading 168 sentences (**item**) that either contained the correct target word, a homophone, or an orthographic control word.

- The surfers traveled to the world-famous *beach/beech/bench* where the waves were very large.

**ffd** = first fixation duration, **sfd** = single fixation duration, **gzd** = gaze duration, **gpt** = go-past time, **tvt** = total time, **skp** = fixation probability (inverse of skipping), **rgi** = regression-in probability, **rgo** = regression-out probability.
```{r read data exp1, echo = FALSE, message=FALSE, warning=FALSE}
## read Exp 1 data and assign approrpiate data types
data <- read.csv("~/Documents/TimeCoursePhonCoding/Revision3_Analyses/SCA.csv")
data$Condition<-ifelse(data$condition==1, "Target", ifelse(data$condition==2, "Homophone", "Control"))
data$Condition<-factor(data$Condition)
data$subject<-factor(data$subject)
data$item<-factor(data$item)
data$condition<-factor(data$condition)
data<-select(data,1:3,5:12,22)
str(data)
head(data)
```
### Means & Standard Errors for Exp. 1

```{r summary statistics exp1, echo = TRUE, message=FALSE, warning=FALSE}
byCond <- group_by(data, Condition)
stats.m<-summarize_if(byCond,is.numeric,funs(mean), na.rm=TRUE)
stats.se<-summarize_if(byCond,is.numeric,funs(std.error), na.rm=TRUE)
stats.m
stats.se
```

### Figure of early measures for Exp. 1
Because it's easier to visualize on a figure than in a table, here are the earliest measures where you can see that readers spent the least amount of time fixating (i.e., processing) the correct target, the most amount of time fixating the orthographic control word, and an intermediate amount of time fixating the homophone--demonstrating an advantage for processing a phonologically related word over a word that shares the same number of letters (i.e., is physically similar).

```{r descriptives plot exp1, echo = FALSE, message=FALSE, warning=FALSE}
data$Condition<- factor(data$Condition,levels(data$Condition)[c(3,2,1)])
dat.SCA<-melt(data=data, id = c("subject","item","Condition"), measure = c("ffd","sfd","gzd"))
p.meanse.SCA<-ggplot(data = subset(dat.SCA, variable %in% c("ffd","sfd","gzd")), aes(y = value, x= Condition, shape=Condition, color=Condition)) + 
  stat_summary(fun.data = "mean_se") + stat_summary(fun.y = mean, geom = "point") + facet_grid(.~variable) +
  labs(y = "Reading Time (ms)", x = "", shape="Condition", color="Condition") + theme_grey(base_size=16) + scale_y_continuous(breaks=seq(0,280,10)) + 
  theme(axis.text.x = element_text(colour="grey4", size=0), axis.text.y = element_text(colour = "grey4", size=20))
p.meanse.SCA + scale_colour_manual(values=c("aquamarine4","aquamarine3","coral3")) + scale_shape_manual(values=c(1,0,2)) 
```

### Survival Analyses for Exp. 1
Although this graded processing suggests that readers activate phonological codes which help them more easily process words which are phonologically related to the expected/contextually appropriate target, just how rapidly these codes come online and begin to influence behavior is not clear. In order to better characterized the time course of code generation, survival analyses were conducted to determine the earliest observable influence of phonology on behavior.

In these analyses, the rate of "surviving" fixations is compared across two condition (here the phonologically related and orthographic control condition) to determine the point at which the survival curves diverge (e.g., the point in time at which more readers have left the phonologically related word than the orthographic control word) which is indicative of easier processing in one condition relative to the other. The divergence point then reflects the earliest observable influence of phonological coding on behavior.

For the paper, I used Matlab to calculate divergence point estimates, however here I am using the new [RTsurvival package](https://github.com/matsukik/RTsurvival)

```{r survdata setup exp1, echo = TRUE, message=FALSE, warning=FALSE}
tmp<-select(data,subject,ffd,condition) %>% arrange(condition) %>%
  rename(duration=ffd)
tmp$condition <- as.numeric(tmp$condition)
survdata<- as.data.frame(tmp %>% as_tibble() %>% mutate(condition = condition-1)) %>%
  filter(condition != 0, !is.na(duration))
survdata$condition <- as.factor(survdata$condition)
str(survdata)
```
We have 48 participants who each have between 35 and 100 data points:

```{r survdata summary Exp 1, echo = TRUE, message=FALSE, warning=FALSE} 
n.per.sbj <- table(survdata$subject)
length(n.per.sbj)
range(n.per.sbj)
```
We can now use these data to generate divergence point estimates (DPE) for each participant:

```{r ip.DPA Exp 1, echo = TRUE, message=FALSE, warning=FALSE}
ip.dpa <- DPA.ip(survdata$subject, survdata$duration, survdata$condition, quiet = TRUE)
dpe <- as.data.frame(ip.dpa$dp_matrix)
# critical columns in output
# 'dpcount' = the number of iterations (out of 1000) on which a DPE was obtained
#' median_dp_duration' = median of the DPEs obtained on each iteration
str(dpe)
dpe$subject[dpe$dpcount<500]
```
Doing so reveals that the DPE for 7 participants were unreliable (i.e., a DP was found on fewer than half of the iterations). Removing those participants reveals a mean DPE of ~173 across the remaining participants (the value moves around ever so slightly each time the bootstrap re-sampling procedure runs). This tells us that on average, phonological coding was influencing behavior by as early as 173 ms after fixation on the target word began.

```{r DPE estimate Exp 1, echo = TRUE, message=FALSE, warning=FALSE}
dpe.rel <- filter(dpe, dpcount >= 500)
summarize(dpe.rel, mean.dpe = mean(median_duration, na.rm=TRUE))
```
Finally, we can represent this visually by examining the survival curves created using the ggsurv function. **Note**--the values displayed on this figure are from the published version of the manuscript and might vary by ~ 1ms from those generated here.

```{r ggsurv function, echo = FALSE, message=FALSE, warning=FALSE}
ggsurv <- function(s, CI = 'def', plot.cens = T, surv.col = c("coral3","aquamarine3"),
                   cens.col = 'red', lty.est = c(2,1), lty.ci = 2,
                   cens.shape = 3, back.white = F, xlab = 'Time',
                   ylab = 'Survival', main = ''){
  
  library(ggplot2)
  strata <- ifelse(is.null(s$strata) ==T, 1, length(s$strata))
  stopifnot(length(surv.col) == 1 | length(surv.col) == strata)
  stopifnot(length(lty.est) == 1 | length(lty.est) == strata)
  
  ggsurv.s <- function(s, CI = 'def', plot.cens = T, surv.col = c("aquamarine4","aquamarine3","coral3"),
                       cens.col = 'red', lty.est = 1, lty.ci = 2,
                       cens.shape = 3, back.white = F, xlab = 'Time',
                       ylab = 'Survival', main = ''){
    
    dat <- data.frame(time = c(0, s$time),
                      surv = c(1, s$surv),
                      up = c(1, s$upper),
                      low = c(1, s$lower),
                      cens = c(0, s$n.censor))
    dat.cens <- subset(dat, cens != 0)
    
    col <- ifelse(surv.col == 'gg.def', 'black', surv.col)
    
    pl <- ggplot(dat, aes(x = time, y = surv)) +
      xlab(xlab) + ylab(ylab) + ggtitle(main) +
      geom_step(col = col, lty = lty.est)
    
    pl <- if(CI == T | CI == 'def') {
      pl + geom_step(aes(y = up), color = col, lty = lty.ci) +
        geom_step(aes(y = low), color = col, lty = lty.ci)
    } else (pl)
    
    pl <- if(plot.cens == T & length(dat.cens) > 0){
      pl + geom_point(data = dat.cens, aes(y = surv), shape = cens.shape,
                      col = cens.col)
    } else if (plot.cens == T & length(dat.cens) == 0){
      stop ('There are no censored observations')
    } else(pl)
    
    pl <- if(back.white == T) {pl + theme_bw()
    } else (pl)
    pl
  }
  
  ggsurv.m <- function(s, CI = 'def', plot.cens = T, surv.col = c("aquamarine4","aquamarine3","coral3"),
                       cens.col = 'red', lty.est = 1, lty.ci = 2,
                       cens.shape = 3, back.white = F, xlab = 'Time',
                       ylab = 'Survival', main = '') {
    n <- s$strata
    
    groups <- factor(unlist(strsplit(names
                                     (s$strata), '='))[seq(2, 2*strata, by = 2)])
    gr.name <-  unlist(strsplit(names(s$strata), '='))[1]
    gr.df <- vector('list', strata)
    ind <- vector('list', strata)
    n.ind <- c(0,n); n.ind <- cumsum(n.ind)
    for(i in 1:strata) ind[[i]] <- (n.ind[i]+1):n.ind[i+1]
    
    for(i in 1:strata){
      gr.df[[i]] <- data.frame(
        time = c(0, s$time[ ind[[i]] ]),
        surv = c(1, s$surv[ ind[[i]] ]),
        up = c(1, s$upper[ ind[[i]] ]),
        low = c(1, s$lower[ ind[[i]] ]),
        cens = c(0, s$n.censor[ ind[[i]] ]),
        group = rep(groups[i], n[i] + 1))
    }
    
    dat <- do.call(rbind, gr.df)
    dat.cens <- subset(dat, cens != 0)
    
    pl <- ggplot(dat, aes(x = time, y = surv, group = group)) +
      xlab(xlab) + ylab(ylab) + ggtitle(main) +
      geom_step(aes(col = group, lty = group))
    
    col <- if(length(surv.col == 1)){
      scale_colour_manual(name = gr.name, values = rep(surv.col, strata))
    } else{
      scale_colour_manual(name = gr.name, values = surv.col)
    }
    
    pl <- if(surv.col[1] != 'gg.def'){
      pl + col
    } else {pl + scale_colour_discrete(name = gr.name)}
    
    line <- if(length(lty.est) == 1){
      scale_linetype_manual(name = gr.name, values = rep(lty.est, strata))
    } else {scale_linetype_manual(name = gr.name, values = lty.est)}
    
    pl <- pl + line
    
    pl <- if(CI == T) {
      if(length(surv.col) > 1 && length(lty.est) > 1){
        stop('Either surv.col or lty.est should be of length 1 in order
             to plot 95% CI with multiple strata')
      }else if((length(surv.col) > 1 | surv.col == c("aquamarine4","aquamarine3","coral3"))[1]){
        pl + geom_step(aes(y = up, color = group), lty = lty.ci) +
          geom_step(aes(y = low, color = group), lty = lty.ci)
      } else{pl +  geom_step(aes(y = up, lty = group), col = surv.col) +
          geom_step(aes(y = low,lty = group), col = surv.col)}
    } else {pl}
    
    
    pl <- if(plot.cens == T & length(dat.cens) > 0){
      pl + geom_point(data = dat.cens, aes(y = surv), shape = cens.shape,
                      col = cens.col)
    } else if (plot.cens == T & length(dat.cens) == 0){
      stop ('There are no censored observations')
    } else(pl)
    
    pl <- if(back.white == T) {pl + theme_bw()
    } else (pl)
    pl
  }
  pl <- if(strata == 1) {ggsurv.s(s, CI , plot.cens, surv.col ,
                                  cens.col, lty.est, lty.ci,
                                  cens.shape, back.white, xlab,
                                  ylab, main)
  } else {ggsurv.m(s, CI, plot.cens, surv.col ,
                   cens.col, lty.est, lty.ci,
                   cens.shape, back.white, xlab,
                   ylab, main)}
  pl
}

```
```{r survival figure Exp 1, echo = TRUE, message=FALSE, warning=FALSE}
data$survdat<-as.integer(ifelse(data$Condition=="Target",NA,data$ffd))
tmp2 <- filter(data, !subject %in% c(19, 20, 21, 25, 27, 30, 38))
ffd.surv <- survfit(Surv(survdat) ~ Condition, data=tmp2)
pl2<-ggsurv(s=ffd.surv)

pl2 + geom_vline(xintercept = 172.61, linetype = "dotted") + 
  annotate("rect", xmin=152.4, xmax=192.9, ymin=0, ymax=1, alpha = .2) + 
  annotate("text", x = 450, y = 0.75, label = "Divergence Point = 173ms", size = 5) +
  annotate("text", x = 450, y = 0.7, label = "95% CI: 152 - 193ms", size = 5) + 
  theme(axis.text.x = element_text(colour="grey4", size=16), axis.text.y = element_text(colour = "grey4", size=16)) + 
  labs(y = "Survival", x = "Time") + theme_grey(base_size=16)
```


## Experiment 2
Data for 48 participants (**subject**) reading 180 sentences (**item**) that either contained the correct target word, a pseudohomophone (i.e., phonologically related non-word), or an orthographic control non-word.

- Because of his insomnia, Caleb couldn’t *sleep/sleap/slerp* even though he was tired.

**ffd** = first fixation duration, **sfd** = single fixation duration, **gzd** = gaze duration, **gpt** = go-past time, **tvt** = total time, **skp** = fixation probability (inverse of skipping), **rgi** = regression-in probability, **rgo** = regression-out probability.
```{r read data exp2, echo = FALSE, message=FALSE, warning=FALSE}
## read Exp 2 data and assign approrpiate data types
data <- read.csv("~/Documents/TimeCoursePhonCoding/Revision3_Analyses/SCB.csv")
data$Condition<-ifelse(data$condition==1, "Target", ifelse(data$condition==2, "Pseudohomophone", "Control"))
data$Condition<-factor(data$Condition)
data$subject<-factor(data$subject)
data$item<-factor(data$item)
data$condition<-factor(data$condition)
data<-select(data,1:3,5:12,18)
str(data)
head(data)
```
### Means & Standard Errors for Exp. 2

```{r summary statistics exp2, echo = TRUE, message=FALSE, warning=FALSE}
byCond <- group_by(data, Condition)
stats.m <- summarize_if(byCond,is.numeric,funs(mean), na.rm=TRUE)
stats.se <- summarize_if(byCond,is.numeric,funs(std.error), na.rm=TRUE)
stats.m
stats.se
```
### Figure of early measures for Exp. 2
As with Experiment 1, the graded processing is clearly visible, where readers have an advantage to processing a phonologically related word over a control word matched on visual similarity. 

```{r descriptives plot exp2, echo = FALSE, message=FALSE, warning=FALSE}
data$Condition<- factor(data$Condition,levels(data$Condition)[c(3,2,1)])
dat.SCA<-melt(data=data, id = c("subject","item","Condition"), measure = c("ffd","sfd","gzd"))
p.meanse.SCA<-ggplot(data = subset(dat.SCA, variable %in% c("ffd","sfd","gzd")), aes(y = value, x= Condition, shape=Condition, color=Condition)) + 
  stat_summary(fun.data = "mean_se") + stat_summary(fun.y = mean, geom = "point") + facet_grid(.~variable) +
  labs(y = "Reading Time (ms)", x = "", shape="Condition", color="Condition") + theme_grey(base_size=16) + scale_y_continuous(breaks=seq(0,310,10)) + 
  theme(axis.text.x = element_text(colour="grey4", size=0), axis.text.y = element_text(colour = "grey4", size=20))
p.meanse.SCA + scale_colour_manual(values=c("aquamarine4","aquamarine3","coral3")) + scale_shape_manual(values=c(1,0,2)) 
```

### Survival Analyses for Exp. 2
Survival analyses were again conducted to determine just how rapidly phonological codes come online and begin to influence behavior, The DPE reflects the earliest observable influence of phonological coding on behavior.

```{r survdata setup exp 2, echo = TRUE, message=FALSE, warning=FALSE}
tmp<-select(data,subject,ffd,condition) %>% arrange(condition) %>%
  rename(duration=ffd)
tmp$condition <- as.numeric(tmp$condition)
survdata<- as.data.frame(tmp %>% as_tibble() %>% mutate(condition = condition-1)) %>%
  filter(condition != 0, !is.na(duration))
survdata$condition <- as.factor(survdata$condition)
str(survdata)
```
We have 48 participants who each have between 55 and 110 data points:

```{r survdata summary Exp 2, echo = TRUE, message=FALSE, warning=FALSE} 
n.per.sbj <- table(survdata$subject)
length(n.per.sbj)
range(n.per.sbj)
```
We can now use these data to generate divergence point estimates (DPE) for each participant:

```{r ip.DPA Exp 2, echo = TRUE, message=FALSE, warning=FALSE}
ip.dpa <- DPA.ip(survdata$subject, survdata$duration, survdata$condition, quiet = TRUE)
dpe <- as.data.frame(ip.dpa$dp_matrix)
# critical columns in output
# 'dpcount' = the number of iterations (out of 1000) on which a DPE was obtained
#' median_dp_duration' = median of the DPEs obtained on each iteration
str(dpe)
dpe$subject[dpe$dpcount<500]
```
Doing so reveals that the DPE for 5 participants were unreliable (i.e., a DP was found on fewer than half of the iterations). Removing those participants reveals a mean DPE of ~161 across the remaining participants (the value moves around ever so slightly each time the bootstrap re-sampling procedure runs). This tells us that on average, phonological coding was influencing behavior by as early as 161 ms after fixation on the target word began.

```{r DPE estimate Exp 2, echo = TRUE, message=FALSE, warning=FALSE}
dpe.rel <- filter(dpe, dpcount >= 500)
summarize(dpe.rel, mean.dpe = mean(median_duration, na.rm=TRUE))
```
Finally, we can represent this visually by examining the survival curves created using the ggsurv function. **Note**--the values displayed on this figure are from the published version of the manuscript and might vary by ~ 1ms from those generated here.

```{r survival figure Exp 2, echo = TRUE, message=FALSE, warning=FALSE}
data$survdat<-as.integer(ifelse(data$Condition=="Target",NA,data$ffd))
tmp2 <- filter(data, !subject %in% c(7, 16, 26, 45, 46))
ffd.surv <- survfit(Surv(survdat) ~ Condition, data=tmp2)
pl2<-ggsurv(s=ffd.surv)

pl2 + geom_vline(xintercept = 161.06, linetype = "dotted") + 
  annotate("rect", xmin=146, xmax=176, ymin=0, ymax=1, alpha = .2) + 
  annotate("text", x = 450, y = 0.75, label = "Divergence Point = 161ms", size = 5) +
  annotate("text", x = 450, y = 0.7, label = "95% CI: 146 - 176ms", size = 5) + 
  theme(axis.text.x = element_text(colour="grey4", size=16), axis.text.y = element_text(colour = "grey4", size=16)) + 
  labs(y = "Survival", x = "Time") + theme_grey(base_size=16)
```

## Experiment 3
Data for 48 participants (**subject**) reading the same 168 sentences (**item**) from Experiment 1. Instead of directly fixating the different target words, all readers fixated the correct target word but received a parafoveal preview that was either identical, the homophone, or the orthographic control word. Previews were presented using the gaze-contingent invisible boundary paradigm (Rayner, 1975).

**ffd** = first fixation duration, **sfd** = single fixation duration, **gzd** = gaze duration, **gpt** = go-past time, **tvt** = total time, **skp** = fixation probability (inverse of skipping), **rgi** = regression-in probability, **rgo** = regression-out probability.
```{r read data exp3, echo = FALSE, message=FALSE, warning=FALSE}
## read Exp 3 data and assign approrpiate data types
data <- read.csv("~/Documents/TimeCoursePhonCoding/Revision3_Analyses/DCA.csv")
data$Condition<-ifelse(data$condition==1, "Target", ifelse(data$condition==2, "Homophone", "Control"))
data$Condition<-factor(data$Condition)
data$subject<-factor(data$subject)
data$item<-factor(data$item)
data$condition<-factor(data$condition)
data<-select(data,1:12,20)
str(data)
head(data)
```
### Means & Standard Errors for Exp. 3

```{r summary statistics exp3, echo = TRUE, message=FALSE, warning=FALSE}
byCond <- group_by(data, Condition)
stats.m <- summarize_if(byCond,is.numeric,funs(mean), na.rm=TRUE)
stats.se <- summarize_if(byCond,is.numeric,funs(std.error), na.rm=TRUE)
stats.m
stats.se
```
### Figure of early measures for Exp. 3
As with Experiment 1, the graded processing is clearly visible, where readers have an advantage to processing a phonologically related word over a control word matched on visual similarity. This demonstrates that the advantage to processing the phonologically related word was not driven by the somewhat strange task of reading contextually inappropriate words, which may have inflated fixation times, especially in the control condition.

```{r descriptives plot exp3, echo = FALSE, message=FALSE, warning=FALSE}
data$Condition<- factor(data$Condition,levels(data$Condition)[c(3,2,1)])
dat.SCA<-melt(data=data, id = c("subject","item","Condition"), measure = c("ffd","sfd","gzd"))
p.meanse.SCA<-ggplot(data = subset(dat.SCA, variable %in% c("ffd","sfd","gzd")), aes(y = value, x= Condition, shape=Condition, color=Condition)) + 
  stat_summary(fun.data = "mean_se") + stat_summary(fun.y = mean, geom = "point") + facet_grid(.~variable) +
  labs(y = "Reading Time (ms)", x = "", shape="Condition", color="Condition") + theme_grey(base_size=16) + scale_y_continuous(breaks=seq(0,310,10)) + 
  theme(axis.text.x = element_text(colour="grey4", size=0), axis.text.y = element_text(colour = "grey4", size=20))
p.meanse.SCA + scale_colour_manual(values=c("aquamarine4","aquamarine3","coral3")) + scale_shape_manual(values=c(1,0,2)) 
```

### Survival Analyses for Exp. 3
Survival analyses were again conducted to determine just how rapidly phonological codes come online and begin to influence behavior, The DPE reflects the earliest observable influence of phonological coding on behavior.

```{r survdata setup exp 3, echo = TRUE, message=FALSE, warning=FALSE}
tmp<-select(data,subject,ffd,condition) %>% arrange(condition) %>%
  rename(duration=ffd)
tmp$condition <- as.numeric(tmp$condition)
survdata<- as.data.frame(tmp %>% as_tibble() %>% mutate(condition = condition-1)) %>%
  filter(condition != 0, !is.na(duration))
survdata$condition <- as.factor(survdata$condition)
str(survdata)
```
We have 48 participants who each have between 33 and 84 data points:

```{r survdata summary Exp 3, echo = TRUE, message=FALSE, warning=FALSE} 
n.per.sbj <- table(survdata$subject)
length(n.per.sbj)
range(n.per.sbj)
```
We can now use these data to generate divergence point estimates (DPE) for each participant:

```{r ip.DPA Exp 3, echo = TRUE, message=FALSE, warning=FALSE}
ip.dpa <- DPA.ip(survdata$subject, survdata$duration, survdata$condition, quiet = TRUE)
dpe <- as.data.frame(ip.dpa$dp_matrix)
# critical columns in output
# 'dpcount' = the number of iterations (out of 1000) on which a DPE was obtained
#' median_dp_duration' = median of the DPEs obtained on each iteration
str(dpe)
dpe$subject[dpe$dpcount<500]
```
Doing so reveals that the DPE for 2 participants were unreliable (i.e., a DP was found on fewer than half of the iterations). Removing those participants reveals a mean DPE of ~170 across the remaining participants (the value moves around ever so slightly each time the bootstrap re-sampling procedure runs). This tells us that on average, phonological coding was influencing behavior by as early as 170 ms after fixation on the target word began.

```{r DPE estimate Exp 3, echo = TRUE, message=FALSE, warning=FALSE}
dpe.rel <- filter(dpe, dpcount >= 500)
summarize(dpe.rel, mean.dpe = mean(median_duration, na.rm=TRUE))
```
Finally, we can represent this visually by examining the survival curves created using the ggsurv function. **Note**--the values displayed on this figure are from the published version of the manuscript and might vary by ~ 1ms from those generated here.

```{r survival figure Exp 3, echo = TRUE, message=FALSE, warning=FALSE}
data$survdat<-as.integer(ifelse(data$Condition=="Target",NA,data$ffd))
tmp2 <- filter(data, !subject %in% c(14, 27))
ffd.surv <- survfit(Surv(survdat) ~ Condition, data=tmp2)
pl2<-ggsurv(s=ffd.surv)

pl2 + geom_vline(xintercept = 170.32, linetype = "dotted") + 
  annotate("rect", xmin=155, xmax=186, ymin=0, ymax=1, alpha = .2) + 
  annotate("text", x = 450, y = 0.75, label = "Divergence Point = 170ms", size = 5) +
  annotate("text", x = 450, y = 0.7, label = "95% CI: 155 - 186ms", size = 5) + 
  theme(axis.text.x = element_text(colour="grey4", size=16), axis.text.y = element_text(colour = "grey4", size=16)) + 
  labs(y = "Survival", x = "Time") + theme_grey(base_size=16)
```

## Experiment 4
Data for 48 participants (**subject**) reading the 180 sentences (**item**) from Experiment 2, with the word-type manipulation done using the gaze-contingent boundary paradigm like Experiment 3, such that all readers directly fixated the correct target after having had a preview that was either identical, a pseudohomophone, or an orthographic control non-word.

**ffd** = first fixation duration, **sfd** = single fixation duration, **gzd** = gaze duration, **gpt** = go-past time, **tvt** = total time, **skp** = fixation probability (inverse of skipping), **rgi** = regression-in probability, **rgo** = regression-out probability.
```{r read data exp4, echo = FALSE, message=FALSE, warning=FALSE}
## read Exp 4 data and assign approrpiate data types
data <- read.csv("~/Documents/TimeCoursePhonCoding/Revision3_Analyses/DCB.csv")
data$Condition<-ifelse(data$condition==1, "Target", ifelse(data$condition==2, "Pseudohomophone", "Control"))
data$Condition<-factor(data$Condition)
data$subject<-factor(data$subject)
data$item<-factor(data$item)
data$condition<-factor(data$condition)
data<-select(data,1:12,19)
str(data)
head(data)
```
### Means & Standard Errors for Exp. 4

```{r summary statistics exp4, echo = TRUE, message=FALSE, warning=FALSE}
byCond <- group_by(data, Condition)
stats.m <- summarize_if(byCond,is.numeric,funs(mean), na.rm=TRUE)
stats.se <- summarize_if(byCond,is.numeric,funs(std.error), na.rm=TRUE)
stats.m
stats.se
```
### Figure of early measures for Exp. 4
As with with the previous 4 experiments, we again see the graded processing indicative of an advantage to processing phonologically related previews.

```{r descriptives plot exp4, echo = FALSE, message=FALSE, warning=FALSE}
data$Condition<- factor(data$Condition,levels(data$Condition)[c(3,2,1)])
dat.SCA<-melt(data=data, id = c("subject","item","Condition"), measure = c("ffd","sfd","gzd"))
p.meanse.SCA<-ggplot(data = subset(dat.SCA, variable %in% c("ffd","sfd","gzd")), aes(y = value, x= Condition, shape=Condition, color=Condition)) + 
  stat_summary(fun.data = "mean_se") + stat_summary(fun.y = mean, geom = "point") + facet_grid(.~variable) +
  labs(y = "Reading Time (ms)", x = "", shape="Condition", color="Condition") + theme_grey(base_size=16) + scale_y_continuous(breaks=seq(0,310,10)) + 
  theme(axis.text.x = element_text(colour="grey4", size=0), axis.text.y = element_text(colour = "grey4", size=20))
p.meanse.SCA + scale_colour_manual(values=c("aquamarine4","aquamarine3","coral3")) + scale_shape_manual(values=c(1,0,2)) 
```

### Survival Analyses for Exp. 4
Survival analyses were again conducted to determine just how rapidly phonological codes come online and begin to influence behavior, The DPE reflects the earliest observable influence of phonological coding on behavior.

```{r survdata setup exp 4, echo = TRUE, message=FALSE, warning=FALSE}
tmp<-select(data,subject,ffd,condition) %>% arrange(condition) %>%
  rename(duration=ffd)
tmp$condition <- as.numeric(tmp$condition)
survdata<- as.data.frame(tmp %>% as_tibble() %>% mutate(condition = condition-1)) %>%
  filter(condition != 0, !is.na(duration))
survdata$condition <- as.factor(survdata$condition)
str(survdata)
```
We have 48 participants who each have between 34 and 98 data points:

```{r survdata summary Exp 4, echo = TRUE, message=FALSE, warning=FALSE} 
n.per.sbj <- table(survdata$subject)
length(n.per.sbj)
range(n.per.sbj)
```
We can now use these data to generate divergence point estimates (DPE) for each participant:

```{r ip.DPA Exp 4, echo = TRUE, message=FALSE, warning=FALSE}
ip.dpa <- DPA.ip(survdata$subject, survdata$duration, survdata$condition, quiet = TRUE)
dpe <- as.data.frame(ip.dpa$dp_matrix)
# critical columns in output
# 'dpcount' = the number of iterations (out of 1000) on which a DPE was obtained
#' median_dp_duration' = median of the DPEs obtained on each iteration
str(dpe)
dpe$subject[dpe$dpcount<500]
```
Doing so reveals that the DPE for 8 participants were unreliable (i.e., a DP was found on fewer than half of the iterations). Removing those participants reveals a mean DPE of ~160 across the remaining participants (the value moves around ever so slightly each time the bootstrap re-sampling procedure runs). This tells us that on average, phonological coding was influencing behavior by as early as 160 ms after fixation on the target word began.

```{r DPE estimate Exp 4, echo = TRUE, message=FALSE, warning=FALSE}
dpe.rel <- filter(dpe, dpcount >= 500)
summarize(dpe.rel, mean.dpe = mean(median_duration, na.rm=TRUE))
```
Finally, we can represent this visually by examining the survival curves created using the ggsurv function. **Note**--the values displayed on this figure are from the published version of the manuscript and might vary by ~ 1ms from those generated here.

```{r survival figure Exp 4, echo = TRUE, message=FALSE, warning=FALSE}
data$survdat<-as.integer(ifelse(data$Condition=="Target",NA,data$ffd))
tmp2 <- filter(data, !subject %in% c( 9, 20, 22, 29, 33, 37, 39, 43))
ffd.surv <- survfit(Surv(survdat) ~ Condition, data=tmp2)
pl2<-ggsurv(s=ffd.surv)

pl2 + geom_vline(xintercept = 159.81, linetype = "dotted") + 
  annotate("rect", xmin=147, xmax=172, ymin=0, ymax=1, alpha = .2) + 
  annotate("text", x = 450, y = 0.75, label = "Divergence Point = 160ms", size = 5) +
  annotate("text", x = 450, y = 0.7, label = "95% CI: 147 - 172ms", size = 5) + 
  theme(axis.text.x = element_text(colour="grey4", size=16), axis.text.y = element_text(colour = "grey4", size=16)) + 
  labs(y = "Survival", x = "Time") + theme_grey(base_size=16)
```
